---
title: "general-voter-analysis-nc"
author: "Luciana Perez Uribe"
date: "2024-05-24"
output: html_document
---

#Dependencies

```{r}
#update.packages(ask = FALSE)
install.packages("dplyr")
install.packages("writexl")
library(writexl)
library(dplyr)
```

#Importing primary election voter registration datasets for 2012,2016,2020 and 2024. 

```{r}
nc_voter_db_2024 <- read.delim("./data/history_stats_20240305.txt", sep = "\t")
nc_voter_db_2020 <- read.delim("./data/history_stats_20200303.txt", sep = "\t")
nc_voter_db_2016 <- read.delim("./data/history_stats_20160315.txt", sep = "\t")
nc_voter_db_2012 <- read.delim("./data/history_stats_20120508.txt", sep = "\t")
```

# party cd = how voters were registered 
# voted_party_cd = in what primary voters participated in 

# Below, we're filtering by party registration. So we're first obtaining the unique party_cd values by year. 


## 2024 
#2024 party registration in the primary. 

```{r}
unique_party_cd <- unique(nc_voter_db_2024$party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(nc_voter_db_2024, party_cd == party)
  #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
}
```

#2024 party participation in the primary. 

```{r}
unique_voted_party_cd <- unique(nc_voter_db_2024$voted_party_cd)
for (party in unique_voted_party_cd) {
  filtered_data <- filter(nc_voter_db_2024, voted_party_cd == party)
  #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
} 
```




##2020 
#2020 party registration in the primary. 

```{r}
unique_party_cd <- unique(nc_voter_db_2020$party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(nc_voter_db_2020, party_cd == party)
   #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
}
```


#2020 party participation in the primary. 

```{r}
unique_voted_party_cd <- unique(nc_voter_db_2020$voted_party_cd)
for (party in unique_voted_party_cd) {
  filtered_data <- filter(nc_voter_db_2024, voted_party_cd == party)
  #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
} 
```



##2016
#2016 party registration in the primary. 

```{r}
unique_party_cd <- unique(nc_voter_db_2016$party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(nc_voter_db_2016, party_cd == party)
  #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
}
```


#2016 party participation in the primary. 

```{r}
unique_voted_party_cd <- unique(nc_voter_db_2016$voted_party_cd)
for (party in unique_voted_party_cd) {
  filtered_data <- filter(nc_voter_db_2016, voted_party_cd == party)
  #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
} 
```



##2012 
#2012 party registration in the primary. 

```{r}
unique_party_cd <- unique(nc_voter_db_2012$party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(nc_voter_db_2012, party_cd == party)
  #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
}
```

#2012 party participation in the primary. 

```{r}
unique_voted_party_cd <- unique(nc_voter_db_2012$voted_party_cd)
for (party in unique_voted_party_cd) {
  filtered_data <- filter(nc_voter_db_2012, voted_party_cd == party)
  #assign(paste0("nc_voter_db_2024_", party), filtered_data)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
} 
```



## Unaffiliated voter analysis 
# For voters who were registered as unaffilated, in what primary did they vote in? 


#2024
```{r}
unaffiliated_ncvoter_2024 <- filter(nc_voter_db_2024, party_cd == "UNA")
sum(unaffiliated_ncvoter_2024$total_voters, na.rm = TRUE)
nique_party_cd <- unique(unaffiliated_ncvoter_2024$voted_party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(unaffiliated_ncvoter_2024, voted_party_cd == party)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
  #assign(paste0("unaffiliated_ncvoter_2024_voted_", party), filtered_data)
}
```

#2020
```{r}
unaffiliated_ncvoter_2020 <- filter(nc_voter_db_2020, party_cd == "UNA")
sum(unaffiliated_ncvoter_2020$total_voters, na.rm = TRUE)
nique_party_cd <- unique(unaffiliated_ncvoter_2020$voted_party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(unaffiliated_ncvoter_2020, voted_party_cd == party)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
  #assign(paste0("unaffiliated_ncvoter_2024_voted_", party), filtered_data)
}
```

#2016
```{r}
unaffiliated_ncvoter_2016 <- filter(nc_voter_db_2016, party_cd == "UNA")
sum(unaffiliated_ncvoter_2016$total_voters, na.rm = TRUE)
nique_party_cd <- unique(unaffiliated_ncvoter_2016$voted_party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(unaffiliated_ncvoter_2016, voted_party_cd == party)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
  #assign(paste0("unaffiliated_ncvoter_2024_voted_", party), filtered_data)
}
```

#2012
```{r}
unaffiliated_ncvoter_2012 <- filter(nc_voter_db_2012, party_cd == "UNA")
sum(unaffiliated_ncvoter_2012$total_voters, na.rm = TRUE)
nique_party_cd <- unique(unaffiliated_ncvoter_2012$voted_party_cd)
for (party in unique_party_cd) {
  filtered_data <- filter(unaffiliated_ncvoter_2012, voted_party_cd == party)
  cat("Party:", party, "- Total Voters:", sum(filtered_data$total_voters, na.rm = TRUE), "\n")
  #assign(paste0("unaffiliated_ncvoter_2024_voted_", party), filtered_data)
}
```

